---
title: "Overview on Functions in R-package kwb.dwd"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Overview}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)

hyperlink <- function(x) sprintf("[%s](%s)", x, x)
```

## List and Download Data from Deutscher Wetterdienst (DWD)

DWD provides data for download via a web interface. The base address to the web interface is: `r hyperlink("https://opendata.dwd.de")`.

There is also an FTP server. The base address to the ftp server
is: `r hyperlink("ftp://opendata.dwd.de")`.

The most relevant files are in
`r hyperlink("ftp://opendata.dwd.de/climate_environment/CDC")`.

You may use the function `ftp_path_cdc` to get this path:

```{r}
kwb.dwd:::ftp_path_cdc()
```

### List Files or Folders in a FTP Directory

The package contains a function to list all files below a certain
FTP address.

Function: `list_url()`

Example:

```{r}
base_url <- kwb.dwd:::ftp_path_cdc("grids_germany")
files <- kwb.dwd::list_url(base_url, max_depth = 1L)
head(files)
```

Set the argument `full_names` to `TRUE` to get the full addresses.

### What files are available on DWD's FTP Server?

Getting a full list of all available files using the `list_url` function takes a while.
Therefore this package contains a data frame with all file paths
as they were available when the package was last updated.

Dataset: `dwd_files`

Example:

```{r}
file_info <- kwb.dwd::dwd_files

# Which files have "evapo" (evaporation) in their name?
is_evapo <- grep("evapo", file_info$file)

head(file_info[is_evapo, c("file", "size", "modification_time")])
```

### Get URLs to Radolan Files

Function: `get_radolan_urls()`

```{r}
radolan_urls <- kwb.dwd::get_radolan_urls()

head(radolan_urls$daily_historical_urls)
tail(radolan_urls$hourly_historical_urls)
```

### Get URLs to Monthly Grids in Zipped ESRI-ascii-grid Format

Function: `list_monthly_grids_germany_asc_gz()`

Example:

```{r eval = FALSE}
urls <- kwb.dwd:::list_monthly_grids_germany_asc_gz("precipitation")

head(urls)
```

```
[1] "ftp://opendata.dwd.de/climate_environment/CDC/grids_germany/monthly/precipitation/01_Jan/grids_germany_monthly_precipitation_188101.asc.gz"
[2] "ftp://opendata.dwd.de/climate_environment/CDC/grids_germany/monthly/precipitation/01_Jan/grids_germany_monthly_precipitation_188201.asc.gz"
[3] "ftp://opendata.dwd.de/climate_environment/CDC/grids_germany/monthly/precipitation/01_Jan/grids_germany_monthly_precipitation_188301.asc.gz"
[4] "ftp://opendata.dwd.de/climate_environment/CDC/grids_germany/monthly/precipitation/01_Jan/grids_germany_monthly_precipitation_188401.asc.gz"
[5] "ftp://opendata.dwd.de/climate_environment/CDC/grids_germany/monthly/precipitation/01_Jan/grids_germany_monthly_precipitation_188501.asc.gz"
[6] "ftp://opendata.dwd.de/climate_environment/CDC/grids_germany/monthly/precipitation/01_Jan/grids_germany_monthly_precipitation_188601.asc.gz"
```

### Get URLs to DWD Metadata

Function: `get_dwd_urls_metadata()`
    
Example:

```{r}
urls_meta <- kwb.dwd::get_dwd_urls_metadata()

head(urls_meta)

# Read the content of one metadata file
writeLines(kwb.utils::readLinesWithEncoding(
  urls_meta[1L], 
  fileEncoding = "WINDOWS-1252", 
  n = 10
))
```

### For a Given URL, Try to Find Description Files

Function: `find_description_files()`

Once you have a URL to a data file, you can look for description files that may 
be contained in the same folder:

```{r}
example_url <- paste0(
  "ftp://opendata.dwd.de/climate_environment/CDC/observations_germany/",    
  "climate/daily/kl/historical/tageswerte_KL_05424_19490301_20211231_hist.zip"
)

kwb.dwd:::find_description_files(example_url)
```

### Open Description for DWD Record in Web Browser

Function: `open_description()`

Example:

```{r eval = FALSE}
kwb.dwd::open_description(example_url)
```

You will be asked something like this if there are description files available:

```
Which file should I open? (0 = Cancel) 

1: BESCHREIBUNG_obsgermany_climate_daily_kl_historical_de.pdf
2: DESCRIPTION_obsgermany_climate_daily_kl_historical_en.pdf
3: KL_Tageswerte_Beschreibung_Stationen.txt
4: KL_Tageswerte_mn4_Beschreibung_Stationen.txt

Auswahl: 
```

### Download Radolan Data in Binary Format from DWD Server

Function: `download_radolan()`

Example:

```{r eval = FALSE}
# Take care, this takes time!
kwb.dwd::download_radolan(resolution = "daily")
```

```{r}
head(dir("~/Downloads/radolan/daily/historical/"))
```

## Download data from other sources

### Download Shape Files for Germany

Function: `check_or_download_shapes_germany()`

This function downloads shape files for Germany. They are not provided by DWD
so this package may not be the best place for this function.

```{r}
target_path <- kwb.dwd::check_or_download_shapes_germany()
```

Shape files have been downloaded and extracted to the target path:

```{r}
shape_files <- dir(target_path, "[.]shp$", recursive = TRUE, full.names = TRUE)
shape_files
```

What can we do with these shape files? We can read and show them, for example:

```{r fig.width = 6}
shape <- sf::read_sf(shape_files[2L])
plot(shape[, "NAME_1"])
```

### Provide List of SpatialPolygonsDataFrame for Germany

Function: `get_shapes_of_germany()`

Example:

```{r eval = FALSE}
shapes <- kwb.dwd::get_shapes_of_germany()
```

### Interactively Configure Selection of SpatialPolygons

Function: `select_shapes()`

Example:

```{r eval = FALSE}
kwb.dwd:::select_shapes(shapes)
```

You will be asked to select a layer...

```
Select layer 

1: gadm40_DEU_1.shp
2: gadm40_DEU_2.shp
3: gadm40_DEU_3.shp
4: gadm40_DEU_4.shp

Auswahl: 
```

### Let User Select a Column from a SpatialPolygonsDataFrame

Function: `select_variable()`

Example:

```{r eval = FALSE}
kwb.dwd:::select_variable(shapes[[2L]])
```

You will be asked to select a variable...

```
Select variable 

1: ID_0 (DEU)
2: COUNTRY (Germany)
3: NAME_1 (Baden-Württemberg, Bayern, Berlin)
4: ID_2 (DEU.1.1_1, DEU.1.2_1, DEU.1.3_1)
5: NAME_2 (Alb-Donau-Kreis, Baden-Baden, Biberach)
6: TYPE_2 (Landkreis, Stadtkreis, Water body)
7: ENGTYPE_2 (District, Water body)
8: CC_2 (08425, 08211, 08426)
9: HASC_2 (DE.BW.AD, DE.BW.BB, DE.BW.BR)
```

## Extract downloaded files

### Extract radolan zip files

Function: `extract_radolan_zip_files()`

### Download a .gz File and Unzip it

Function: `unzip_asc_gz_file()`

## Conversion

### Convert Binary Radolan File to Raster File

Function: `convert_bin_to_raster_file()`

We need a binary file as provided by DWD. See above for how to download 
Radolan binary files.

```{r}
bin_file <- "~/Downloads/radolan/raa01-sf_10000-1808041150-dwd---bin"
raster <- kwb.dwd::convert_bin_to_raster_file(bin_file)
```

We can plot the content of the raster file using the `plot` function of the 
raster package:

```{r fig.width = 6}
raster::plot(raster)
```

### Radolan raw to raster

Function: `radolan_raw_to_raster()`

### Conversion of coordinates

#### Coordinates to EPSG-4326

Function: `coordinates_to_EPSG_4326()`

```{r}
kwb.dwd::coordinates_to_EPSG_4326(longitude = 0, latitude = 0)
```

Where do we need this? Good question, next question.

### Cropping

Crop a circle-shaped region from the Radolan data grid of Germany

Function: `crop_radolan_radial()`

We need a folder with rst files...

```{r eval = FALSE}
rst_path <- "~/Downloads/rst"

files <- dir(rst_path, full.names = TRUE)

radolan_data <- raster::stack(files)

names(radolan_data) <- substr(
  gsub("raa01-sf_10000-", "", basename(files)), 
  start = 1, 
  stop = 6
)
  
# Error!
#radolan_data <- kwb.dwd::read_relevant_years_radolan(
#  rst_path, years = 2018
#)

cropped <- kwb.dwd::crop_radolan_radial(
  radolan_data, 
  latitude = 52.520008, 
  longitude = 13.404954
)

# What does the warning mean?
cropped
```

**Problem here:** `read_relevant_years_radolan` is too specific, expects a certain file name pattern 

## Functions Related to Radolan Data

that are not alread mentioned elsewhere...

### Get Radolan Extension ("Bounding box" of Radolan area)

Function: `get_radolan_extension()`

```{r}
kwb.dwd::get_radolan_extension()
```

**Which coordinate system?**

### Compose radolan projection string

Function: `get_radolan_projection_string()`

```{r}
kwb.dwd::get_radolan_projection_string()
```

## Read different file formats

### Read Zipped ESRI-Ascii-Grid File (from URL)

Function: `read_asc_gz_file()`

### Read an Example Raster File for Germany

Function: `get_example_grid_germany()`

Example:

```{r}
grid <- kwb.dwd::get_example_grid_germany()
raster::plot(grid)
```

### Read Binary Radolan File

Function: `read_binary_radolan_file()`

### Read Relevant Radolan Data from rst Database

Function: `read_relevant_years_radolan()`

## Write different file formats

### Writing raster data to a file

Function: `write_raster_to_file()`

```{r}
kwb.dwd::write_raster_to_file(
  raster, 
  file.path(tempdir(), "raster.rst")
)
```

## High-Level Functions, General

### Read daily data from DWD, mask region with given shape file

Function: `read_daily_data_over_shape()`

### Read monthly data from DWD, mask region with given shape file

Function: `read_monthly_data_over_shape()`
    
### Download, crop, summarise

Function: `calculate_masked_grid_stats()`

This seems to be a highly specific function that requires a list of matrices as input. But where to get this list from... ? We can create an example from 
scratch to demonstrate what the function does:

```{r}
#writeLines(paste(sample(0:3, 12, replace = TRUE), collapse = ", "))
to_matrix <- function(...) matrix(c(...), ncol = 3, byrow = TRUE)

matrix_1 <- structure(file = "file_1", year = 2022L, month = 1L,  to_matrix(
  2, 3, 1, 
  0, 2, 3, 
  1, 3, 2, 
  1, 2, 0
))

matrix_2 <- structure(file = "file_2", year = 2022L, month = 2L, to_matrix(
  3, 0, 2, 
  0, 0, 0, 
  3, 0, 1, 
  3, 2, 1
))

matrices <- list(matrix_1, matrix_2)

# Take car to use NA, not zero!
geo_mask <- to_matrix(
  NA, 1, NA, 
   1, 1,  1, 
   1, 1,  1, 
  NA, 1, NA
)

kwb.dwd::calculate_masked_grid_stats(matrices, geo_mask)
```

This was a simple example. We have used this function to calculate mean 
evaporation values for Berlin. Thanks to Andreas Matzinger we have a function 
that returns the matrix that can be used to crop the Berlin area.

Function: `get_berlin_dwd_mask`

```{r}
raster::plot(raster::raster(kwb.dwd::get_berlin_dwd_mask()))
```

## High-Level Functions, Specific to Berlin

### Load monthly potential evaporation for Berlin from DWD

Function: `load_potential_evaporation_berlin()`
Function: `load_potential_evaporation_berlin_2()` (newer version)
    
### Load monthly precipitation for Berlin from DWD

Function: `load_precipitation_berlin()`
